---
title: "Predicting Stock Returns with Cluster-Then-Predict"
author: "Terrel Shumway"
date: "04/20/2015"
output: html_document
---
  
This presents the answers for questions in part two of assignment 6.

```{r}
baseurl = "https://courses.edx.org/c4x/MITx/15.071x_2/asset/"

getdata = function(local){
  if(!file.exists(local)){
    library(downloader)
    remote = paste0(baseurl,local)
    download(remote,local)
  }
  read.csv(local,stringsAsFactors=FALSE)
}

data = getdata("StocksCluster.csv")


```

## Exploring the Dataset
problem 1.1: `r nrow(data)`

problem 1.2: `r mean(data$PositiveDec)`

```{r,results='asis'}
knitr::kable(cor(data))
cc = sort(cor(data),decreasing=TRUE)[13]
```
problem 1.3: `r cc`

problem 1.4
```{r}
indep = setdiff(names(data),"PositiveDec")
barplot(colMeans(data[,indep]))
```


## Logistic Regression Model

```{r}
library(caTools)
set.seed(144)
spl = sample.split(data$PositiveDec,SplitRatio=0.7)
train = subset(data,spl==TRUE)
test = subset(data,spl!=TRUE)

mystats = function(conf){
  n = sum(conf)
  tp = conf[2,2]
  tn = conf[1,1]
  fp = conf[2,1]
  fn = conf[1,2]
  acc = (tp+tn)/n
  spec = tp/(tp+fn)
  sens = tn/(tn+fp)
  c(tp=tp,tn=tn,fp=fp,fn=fn,acc=acc,spec=spec,sens=sens)
}

confacc = function(model,data,level=0.5,depvar="PositiveDec"){
  pred = predict(model,newdata=data,method="response")
  conf = table(data[,depvar],pred>level)
  #acc = (conf[1,1]+conf[2,2])/nrow(data)
  print(conf)
  mystats(conf)
}


m1 = glm(PositiveDec ~ ., data=train, family="binomial")
summary(m1)

pred = predict(m1,method="response")
conf = table(train$PositiveDec,pred>0.5)

m1.acc = mystats(conf)

```
problem 2.1: accuracy=`r m1.acc`

```{r}
m1.acc = confacc(m1,test)
```
problem 2.2: accuracy=`r m1.acc`

```{r}

bl.conf = table(test$PositiveDec)
bl.acc = bl.conf[2]/nrow(test)

```
problem 2.3: accuracy=`r bl.acc`

## Cluster then Predict

```{r}
ltrain = train
ltest = test
ltrain$PositiveDec = NULL
ltest$PositiveDec = NULL

library(caret)
preproc = preProcess(ltrain)
ntrain = predict(preproc,ltrain)
ntest = predict(preproc,ltest)
```
problem 3.2: `r mean(ntrain$ReturnJan)` `r mean(ntest$ReturnJan)`


```{r}
set.seed(144)
kluster = kmeans(ntrain,3)
```
problem 3.4: `r which.max(kluster$size)`

problem 3.5:
```{r}
library(flexclust)
storage = "stockcluster.Rdata.xz"
if(file.exists(storage)){
  load(storage)  
}else{  
  km.kcca = as.kcca(kluster,ntrain)
  save(km.kcca,file=storage,compress="xz")
}

ctrain = predict(km.kcca)
ctest = predict(km.kcca,newdata=ntest)
```

problem 4.1
```{r}
table(ctest)
```

problem 4.2

```{r}

sm1 = glm(PositiveDec ~ ., data=train[ctrain==1,],family="binomial")
summary(sm1)
sm2 = glm(PositiveDec ~ ., data=train[ctrain==2,],family="binomial")
summary(sm2)
sm3 = glm(PositiveDec ~ ., data=train[ctrain==3,],family="binomial")
summary(sm3)
names(sm3)
signs = cbind(s1=sign(sm1$coefficients),s2 = sign(sm2$coefficients),s3 = sign(sm3$coefficients))
signs
```


```{r}
ac1 = confacc(sm1,test[ctrain==1,])
ac2 = confacc(sm2,test[ctrain==2,])
ac3 = confacc(sm3,test[ctrain==3,])
c(ac1,ac2,ac3)
```

