---
title: "Trees"
author: "Terrel Shumway"
date: "03/30/2015"
output: html_document
---

First, we download and read the file `ClaimsData.csv`...

```{r}


claims = read.csv("ClaimsData.csv")
str(claims)

```

split the data into Train and Test sets...

```{r}
#install.packages("caTools")
library(caTools)
set.seed(88)
spl = sample.split(claims$bucket2009,SplitRatio=0.6)
Train = subset(claims,spl)
Test = subset(claims,!spl)

sapply(Train[c("age","diabetes")],mean)

e = function( outcome,prediction){
  conf = table(outcome,prediction)
  n = nrow(outcome)
  accuracy = (conf[1,1]+conf[2,2]+conf[3,3]+conf[4,4]+conf[5,5])/n
  PenaltyMatrix = matrix(c(0,1,2,3,4,
                         2,0,1,2,3,
                         4,2,0,1,2,
                         6,4,2,0,1,
                         8,6,4,2,0),
                       byrow=T,nrow=5)

  penalty = sum(as.matrix(conf)*PenaltyMatrix)/n
  print(accuracy)
  print(penalty)
  c(accuracy=accuracy,penalty=penalty)
}

e(Test$bucket2009,Test$bucket2008)

```

## CART

All of our models are going to use the same independent variables. lets set this once.
```{r}
stformula = bucket2009 ~ age + arthritis + alzheimers + 
    cancer + copd + depression + diabetes + heart.failure +
    ihd + kidney + osteoporosis + stroke +
    bucket2008 + reimbursement2008 

library(rpart)
library(rpart.plot)
Tree = rpart( stformula, data=Train, method="class",cp=0.00005,parms=list(loss=PenaltyMatrix))
#prp(Tree)
p1 = predict(Tree,newdata=Test,type="class")
conf=table(Test$bucket2009,p1)
conf
```


Now, we are ready to build our CART model. First we need to install and load the packages rpart and rpart.plot

```{r}
#install.packages("rpart")
library(rpart)
#install.packages("rpart.plot")
library(rpart.plot)



st = rpart( stformula, 
            data=Train,
            method="class", minbucket=25)

prp(st)
```

Now, let's see how well we did:

```{r}
p1 = predict(st, newdata=Test, type="class")
conf = table(Test$Reverse,p1)
accuracy = (conf[1]+conf[4])/nrow(Test)
print(conf)
```

The accuracy is `r accuracy`.

```{r}
library(ROCR)
p2 = predict(st,newdata=Test)
pred = prediction(p2[,2],Test$Reverse)
perf = performance(pred,"tpr","fpr")
plot(perf)
auc = as.numeric(performance(pred,"auc")@y.values)
```

The area under this curve is `r auc`.

If we increase and decrease the `minbucket` parameter

```{r echo=FALSE}
csplits = function(x){x$cptable[nrow(x$cptable),"nsplit"]}

st5 = rpart(stformula, 
            data=Train,
            method="class", minbucket=5)

st100 = rpart(stformula, 
            data=Train,
            method="class", minbucket=100)

```

we get minbucket=5 -> `r csplits(st5)` and minbucket=100 -> `r csplits(st100)` splits.

## Random Forests

```{r}
# install.packages("randomForest")
library(randomForest)

set.seed(100)
stevensForest = randomForest(stformula, 
            data=Train,
            nodesize = 25, ntree = 200)


pf = predict(stevensForest, newdata=Test)
conf = table(Test$Reverse,pf)
accuracy = (conf[1]+conf[4])/nrow(Test)
print(conf)
```

The accuracy is `r accuracy`.

```{r}
set.seed(200)
stevensForest = randomForest(stformula, 
            data=Train,
            nodesize = 25, ntree = 200)


pf = predict(stevensForest, newdata=Test)
conf = table(Test$Reverse,pf)
accuracy = (conf[1]+conf[4])/nrow(Test)
print(conf)
```

The accuracy is `r accuracy`.

## k-Fold Cross Validation

```{r}
#install.packages(c("caret","e1071"))
library(caret)
library(e1071)

numFolds = trainControl(method="cv",number=10)
cpGrid = expand.grid(.cp=seq(0.01,0.5,0.01))
cvmodels = train(stformula, 
      data=Train,
      method="rpart",
      trControl=numFolds,
      tuneGrid = cpGrid)

bestcp = cvmodels$bestTune
```

The best value for tuning the CART model is `r bestcp`. We'll use this create the new CART model.

```{r}
stevensCV = rpart(stformula,data=Train,method="class",cp=bestcp)
prp(stevensCV)
```

This model has `r csplits(stevensCV)` split(s).

```{r}
p1 = predict(stevensCV, newdata=Test, type="class")
conf = table(Test$Reverse,p1)
accuracy = (conf[1]+conf[4])/nrow(Test)
print(conf)
```

The accuracy is `r accuracy`.
