---
title: "Trees"
author: "Terrel Shumway"
date: "03/30/2015"
output: html_document
---

First, we download and read the file `stevens.csv`...

```{r}

baseurl = "https://courses.edx.org/c4x/MITx/15.071x_2/asset/"

get_data = function(url,local){
  if(!file.exists(local)){
    download.file(url,local,"curl")
  }
  read.csv(local)
}

data_dir = function(fname){
  fname
  #paste("data",fname,sep="/")
}

getm = function(file){
  get_data(paste(baseurl,file,sep=""),data_dir(file))
}


stevens = getm("stevens.csv")
stevens$Reverse = as.factor(stevens$Reverse)
str(stevens)

```

split the data into train and test sets...

```{r}
#install.packages("caTools")
library(caTools)
set.seed(3000)
spl = sample.split(stevens$Reverse,SplitRatio=0.7)
train = subset(stevens,spl)
test = subset(stevens,!spl)

```

## CART

Now, we are ready to build our CART model. First we need to install and load the packages rpart and rpart.plot

```{r}
#install.packages("rpart")
library(rpart)
#install.packages("rpart.plot")
library(rpart.plot)

st = rpart( Reverse ~ Circuit + Issue + Petitioner + Respondent + LowerCourt + Unconst, 
            data=train,
            method="class", minbucket=25)

prp(st)
```

Now, let's see how well we did:

```{r}
p1 = predict(st, newdata=test, type="class")
conf = table(test$Reverse,p1)
accuracy = (conf[1]+conf[4])/nrow(test)
print(conf)
```

The accuracy is `r accuracy`.

```{r}
library(ROCR)
p2 = predict(st,newdata=test)
pred = prediction(p2[,2],test$Reverse)
perf = performance(pred,"tpr","fpr")
plot(perf)
auc = as.numeric(performance(pred,"auc")@y.values)
```

The area under this curve is `r auc`.

If we increase and decrease the `minbucket` parameter

```{r echo=FALSE}
csplits = function(x){x$cptable[nrow(x$cptable),"nsplit"]}

st5 = rpart( Reverse ~ Circuit + Issue + Petitioner + Respondent + LowerCourt + Unconst, 
            data=train,
            method="class", minbucket=5)

st100 = rpart( Reverse ~ Circuit + Issue + Petitioner + Respondent + LowerCourt + Unconst, 
            data=train,
            method="class", minbucket=100)

```

we get minbucket=5 -> `r csplits(st5)` and minbucket=100 -> `r csplits(st100)` splits.

## Random Forests

```{r}
# install.packages("randomForest")
library(randomForest)

set.seed(100)
stevensForest = randomForest(Reverse ~ Circuit + Issue + Petitioner + Respondent + LowerCourt + Unconst, 
            data=train,
            nodesize = 25, ntree = 200)


pf = predict(stevensForest, newdata=test)
conf = table(test$Reverse,pf)
accuracy = (conf[1]+conf[4])/nrow(test)
print(conf)
```

The accuracy is `r accuracy`.

```{r}
set.seed(200)
stevensForest = randomForest(Reverse ~ Circuit + Issue + Petitioner + Respondent + LowerCourt + Unconst, 
            data=train,
            nodesize = 25, ntree = 200)


pf = predict(stevensForest, newdata=test)
conf = table(test$Reverse,pf)
accuracy = (conf[1]+conf[4])/nrow(test)
print(conf)
```

The accuracy is `r accuracy`.
