---
title: "Music Hits"
author: "Terrel Shumway"
date: "04/14/2015"
output: html_document
---

Load the data.

```{r}
remote = "https://courses.edx.org/c4x/MITx/15.071x_2/asset/songs.csv"
local = "songs.csv"
if(!file.exists(local)){
  library(downloader)
  download(remote,local)
}
songs = read.csv(local)
```


## Understanding the Data

```{r}
byyear = table(songs$year)
```
Problem 1.1: `r byyear["2010"]`



```{r}
jackson = subset(songs,artistname=="Michael Jackson")
```
Problem 1.2: `r nrow(jackson)`

Problem 1.3:
```{r,results='asis'}
knitr::kable(subset(jackson,jackson$Top10==1)[,c("year","songtitle")])
```


Problem 1.4:
```{r,results='asis'}
tsig = data.frame(sort(table(songs$timesignature)))
knitr::kable(tsig)
```

Problem 1.5: The fastest song is: `r songs[which.max(songs$tempo),"songtitle"]`


## Creating a Prediction Model

```{r}
nonvars = c("year", "songtitle", "artistname", "songID", "artistID")
vars = setdiff(colnames(songs),nonvars)

spl = songs$year <= 2009
train = subset(songs,spl)[,vars]
test = subset(songs,!spl)[,vars]
```
Problem 2.1: There are `r nrow(train)` songs in the training set.

Problems 2.2-2.5 can be answered from the summary below:
```{r}
m1 = glm(Top10~.,data=train,family="binomial")
summary(m1)
```

## Beware of Multicolinearity Issues

```{r}
m2 = glm(Top10~.-loudness,data=train,family="binomial")
summary(m2)
```

```{r}
m3 = glm(Top10~.-energy,data=train,family="binomial")
summary(m3)
```

## Validating Our Model

```{r}
p3 = predict(m3,newdata=test,type="response")
m3.conf = table(test$Top10,p3>.45)
m3.acc = (m3.conf[1,1]+m3.conf[2,2])/nrow(test)
```
Problem 4.1: accuracy = `r m3.acc`

```{r}
baseline = table(test$Top10)
bl.acc = baseline[1]/nrow(test)
```
Problem 4.2: baseline accuracy = `r bl.acc`

Problem 4.3:
```{r,results='asis'}
knitr::kable(m3.conf)
```

Problem 4.4:
```{r}
tn = m3.conf[1,1]
tp = m3.conf[2,2]
fn = m3.conf[2,1]
fp = m3.conf[1,2]
sens = tp/(tp+fn)
spec = tn/(fp+tn)
acc = (tp+tn)/nrow(test)
```

---------------|-------------
true negative  | `r tn`
true positive  | `r tp`
false negative | `r fn`
false positive | `r fp`
accuracy       | `r acc`
sensitivity    | `r sens`
specificity    | `r spec`


This model favors specificity over sensitivity. It is a conservative model that will miss many hits, but provides high confidence that selected songs will become hits.
